A local search approach for flexible job shop problems

### 2016-01-09 paper/paper.tex
updated paper, cooling factor

\documentclass[a4paper,12pt]{article}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{tikz}

@@

The Flexible Job Shop problem is defined as follows.

\subsubsection{Notation}

We are given a set of $n$ jobs.

@@

\subsubsection{Objective}

@@

\subsubsection{Definition considerations}

@@

The classical Job Shop problem is often modeled as a disjunctive graph (for example, see \textbf{Figure 1}), where the directed arcs represent the precedence constraints within a job, and the undirected edges represent operations that must be processed on the same machine. By defining a direction on the edges, and making sure that these definitions result in an acyclic graph, one obtains a valid schedule for the Job Shop instance. In this case, an undirected edge represents a non-simultaneity constraint -- the two operations may be processed in any order, but not at the same time. If we wish to use the same definition for the FSJ, we are forced to assign operations to machines before directing the edges. However, that splits the problem into two parts as mentioned above, and will increase the complexity of the neighborhood. We will take another approach.

\begin{figure}
\centering
\begin{tikzpicture}

@@

\end{tikzpicture}
\caption{A disjunctive graph representing 3 jobs on 3 machines}
\end{figure}

@@

In our representation, we can ascertain the following information from \textbf{Figure 1}. There are three jobs. Operations 1 and 8 must be processed by the same machine type. Similarly, operations 2, 4, and 6 are processed on another, and 3, 5, and 7 on the last machine type.

@@

To do this, we place the unscheduled operations in ordered queues for each machine type $T_k$. We take the first operation that has no unscheduled job predecessor, and schedule it greedily into the first available machine, taking care to schedule it at the same start time or after that of the previously scheduled operation for that type. This conforms to the requirements we laid out earlier in the representation section.

@@

We could choose to apply tricks to generate a better schedule, but instead we will allow the local search to do that by modifying the order of the operations for each machine type.

@@

\begin{figure}
\centering
\begin{tikzpicture}

@@

% edges
\draw[arc] (1) to (8);
\draw[arc] (5) to (3);
\draw[arc] [bend left=20] (7) to (3);
\draw[arc] (7) to (5);
\draw[arc] (2) to (6);
\draw[arc] (6) to (4);

\end{tikzpicture}
\caption{An acyclic directed graph representing a schedule}
\end{figure}

Consider the acyclic directed graph in \textbf{Figure 2}. The serialization function would set up three ordered queues.

@@

A neighborhood solution is generated by calculating the longest path

@@

\bibitem{approxfjsp}
	Jansen, Klaus, Monaldo Mastrolilli, Roberto Solis-Oba.
	``Approximation algorithms for flexible job shop problems.''
	\textit{LATIN 2000: Theoretical Informatics}. Springer Berlin Heidelberg, 2000. 68-77.

>>>>

\documentclass[a4paper,10pt]{article}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{tikz}

@@

The Flexible Job Shop problem is defined as follows. We are given a set of $n$ jobs.

@@

@@

@@

The classical Job Shop problem is often modeled as a disjunctive graph (for example, see \textbf{Figure \ref{fig:disjunctivegraph}}), where the directed arcs represent the precedence constraints within a job, and the undirected edges represent operations that must be processed on the same machine. By defining a direction on the edges, and making sure that these definitions result in an acyclic graph, one obtains a valid schedule for the Job Shop instance. In this case, an undirected edge represents a non-simultaneity constraint -- the two operations may be processed in any order, but not at the same time. If we wish to use the same definition for the FSJ, we are forced to assign operations to machines before directing the edges. However, that splits the problem into two parts as mentioned above, and will increase the complexity of the neighborhood. We will take another approach.

\begin{figure}[h]
\centering
\begin{tikzpicture}

@@

\end{tikzpicture}
\caption{A disjunctive graph representing 3 jobs on 3 machines}
\label{fig:disjunctivegraph}
\end{figure}

@@

In our representation, we can ascertain the following information from \textbf{Figure \ref{fig:disjunctivegraph}}. There are three jobs. Operations 1 and 8 must be processed by the same machine type. Similarly, operations 2, 4, and 6 are processed on another, and 3, 5, and 7 on the last machine type.

@@

To do this, we place the unscheduled operations in ordered queues for each machine type $T_k$. We take the first operation that has no unscheduled job predecessor, and schedule it greedily into the first available machine, taking care to schedule it at the same start time or after that of the previously scheduled operation for that type. This ensures the ordering on the operations is preserved in the schedule, and conforms to the requirements we laid out earlier in the representation section. Scheduling operations before the starting time of a previously scheduled operation on the same machine type can lead to a schedule that is actually infeasible.

@@

We could choose to apply several tricks to generate a better schedule, but instead we will let the local search do that for us by modifying the order of the operations for each machine type.

@@

\begin{figure}[h]
\centering
\begin{tikzpicture}

@@

% edges
\draw[arc] (1) to (8);
\draw[arc] (5) to (3);
\draw[arc] [bend right=20] (7) to (3);
\draw[arc] (7) to (5);
\draw[arc] (2) to (6);
\draw[arc] (6) to (4);

\end{tikzpicture}
\caption{An acyclic directed graph representing a schedule}
\label{fig:dag}
\end{figure}

Consider the acyclic directed graph in \textbf{Figure \ref{fig:dag}}. The serialization function would set up three ordered queues:

@@

For the classical Job Shop problem, a commonly used neighborhood is the set of all critical block swaps. The critical block swap is a reversal of an edge on the longest path of the disjunctive graph. This neighborhood is quite small, focusing on the operations that influence the length of the makespan. More importantly, the neighborhood is guaranteed to consist only of other acyclic graphs, or feasible solutions \cite{simulatedannealingjsp}.

For the FJS, we can define the longest path in a similar manner. Whereas the longest path for the classical Job Shop problem was also chronologically sequential in terms of the generated schedule, this is no longer necessarily the case for the FJS. We must also consider operations on the same machine type, scheduled on different machines. For example, consider the operations $O_{1,1}$ and $O_{2,2}$ from different jobs that must be scheduled on machines of type $T$, connected by a directed edge from $O_{1,1}$ to $O_{2,2}$. Suppose $O_{1,1}$ has been scheduled at time $t_1$ on machine $M_1$ of type $T$ due to a precedence constraint. Further suppose that machine $M_2$, also of type $T$, is available from some time $t_2 < t_1$. However, the serialization function is forced to schedule $O_{2,2}$ on $M_2$ at time $t_1$ anyway. In this scenario, we would like to reverse the order of the arc between $O_{1,1}$ and $O_{2,2}$, so that $O_{2,2}$ may be scheduled earlier and therefore potentially decrease the length of the longest path.

Therefore, we construct our neighborhood by considering operations $v$ and $w$ such that:

\begin{itemize}
	\item $v$ and $w$ have both been scheduled consecutively on the same machine
	\item $v$ and $w$ have been scheduled on different machines of the same type at the same time
\end{itemize}

\subsubsection{Applying the swap}

Let's consider the absolute ordering on the set of operations on a machine type, as defined by the acyclic directed graph corresponding to a feasible schedule. In the classical Job Shop case, a critical block swap identifies two consecutively scheduled nodes on a machine. These two nodes are necessarily consecutive in the absolute ordering, as well. In the FJS problem, this is no longer necessarily the case, since operations in between may have been scheduled on other machines of the same type. If one of these other operations is of the same job as either of the selected nodes, simply reversing the directed edge between the two nodes will lead to one or more cycles.

Consider the following absolute ordering of a machine type. \textbf{Figure \ref{fig:swap}} shows a possible scheduling configuration for this sample, where similar colors indicate the same job.

\[ \dots \rightarrow O_{1,9} \rightarrow O_{2,12} \rightarrow O_{1,10} \rightarrow O_{2,13} \rightarrow \dots \]

We wish to swap the order of operations $O_{1,9}$ and $O_{2,13}$, both on the critical path. We can resolve this situation by reversing a few more arcs, to achieve the following absolute ordering.

\[ \dots \rightarrow O_{2,12} \rightarrow O_{2,13} \rightarrow O_{1,9} \rightarrow O_{1,10} \rightarrow \dots \]

The swap operation must preserve the job order of all operations in the graph, as above. This ensures that no cycles will be created within a machine type.

\subsubsection{Infeasible solutions}

The critical path neighborhood for the classical Job Shop is guaranteed to consist of only feasible schedules, a very nice result from Van Laarhoven et al \cite{simulatedannealingjsp}. Unfortunately, this is not the case for the Flexible Job Shop. Consider the schedule shown in \textbf{Figure \ref{fig:cycle}}, and the associated absolute ordering below:

@@

\bibitem{approxfjsp}
	Jansen, Klaus, Monaldo Mastrolilli, and Roberto Solis-Oba.
	``Approximation algorithms for flexible job shop problems.''
	\textit{LATIN 2000: Theoretical Informatics}.
	Springer Berlin Heidelberg, 2000. 68-77.

\bibitem{simulatedannealingjsp}
	Van Laarhoven, Peter JM, Emile HL Aarts, and Jan Karel Lenstra.
	``Job shop scheduling by simulated annealing."
	\textit{Operations research}.
	40.1 (1992): 113-125.
